---
title: "Vesalius: Quick Start"
author: "Patrick C.N. Martin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Vesalius: Quick Start

Do you like to live in the fast lane? Do you sigh every time you see a long
vignette? Do you wish there was a TLDR on everything?

I understand.

Welcome to the Vesalius quick start!

In this vignette, you will find a **very** short guide on how to work with Vesalius.
Vesalius provides an internal data set taken from real Spatial transcrtiptomic
data. For an in depth view of Vesalius, please refer to Vesalius_Analysis.Rmd.

For all intents and purposes, this is a dummy data set and should only be used
to get a feel for the Vesalius package. We strongly advise you to use real and
complete data sets to fully understand the Vesalius package and its benefits.  

# Introduction
Vesalius is a tool to perform high-resolution in silico anatomization and
molecular characterization from ST data without requiring companion images.

This is achieved by converting ST data into images: PCA loading values are
embedded into RGB colour channels by taking the sum of the absolute value of all
loading values associated to each bead on the ST array.

Image analysis techniques are applied to the image and anatomical territories
can subsequently be extracted.

# Loading

First, let's load the package and the data. Data originates from slide-seqV2
data (availble at (Single Cell Portal)[https://singlecell.broadinstitute.org/single_cell/study/SCP948/robust-decomposition-of-cell-type-mixtures-in-spatial-transcriptomics#study-download]).

For sake of simplicity, Vesalius takes a small subset of this data set and
stores it in the form of a Seurat object.

```{r loading, eval = TRUE, echo = TRUE}
library(vesalius)
data(vesalius,package = "vesalius")
ls()
```

# Pre-processing
Second, we pre-process the data using Seurat

```{r preprocessing, eval = TRUE, echo = TRUE}
# First create a copy of raw data for later use
# When working on isolated territories, we want to run pre-processing
# on raw counts not on already transformed counts.
copy <- vesalius

# Preprocessing Spatial transcriptomic data
vesalius <- NormalizeData(vesalius)
vesalius <- FindVariableFeatures(vesalius, nfeatures = 2000)
vesalius <- ScaleData(vesalius)

```

# RGB to PCA
Third, we convert a count matrix (stored in a Seurat object) to RGB coded
coordinates. We will only produce one slice (i.e. PC1 to PC3)

```{r rgbtopca, eval = TRUE, echo = TRUE}
vesalius <- rgbPCA(vesalius,slices = 1)
```

# Building an Image
Forth, we create an actual image by using a combination of voronoi diagrams and
tile rasterisation.

```{r buildImage,eval = TRUE, echo = TRUE,fig.height = 6, fig.width = 6}
vesalius <- buildImageArray(vesalius, sliceID = 1, invert =T)


# Let's have a look shall we?
Vesalius_image <- imagePlot(vesalius,as.cimg =FALSE) + theme_void()


print(Vesalius_image)

```

# Applying Image Analysis techniques
Firth, we can apply image analysis techniques such as histogram equalisation,
variance regularisation, smoothing and segmentation.

```{r imageAnalysis, eval = TRUE , echo = TRUE}
# Histogram EQ
vesalius <- equalizeHistogram(vesalius,sleft = 2.5, sright=2.5,invert =T)


# Variance regularisation
vesalius <- regulariseImage(vesalius, lambda = 10,
                            niter = 200, normalise=T)


# Smoothing
# !!!Optional!!! -
vesalius <- smoothArray(vesalius,method = "box", box = 1)

# Segmentation
# This function provides smoothing internally.
vesalius <- iterativeSegmentation.array(vesalius,
                                        colDepth = 6,
                                        smoothIter = 20,
                                        method = c("iso","median"),
                                        sigma=1.5,
                                        box = 10,
                                        useCenter = T,
                                        invert =T)


# Isolating territories from colour segments
vesalius <- isolateTerritories.array(vesalius,
                                     captureRadius = 0.1,
                                     minBar = 1)

vesalius

```

# Visualisation
Sixth, we can visualise the territories.

```{r viz, eval = TRUE, echo = TRUE, fig.width = 6, fig.height = 6}
# Let's have a look!
imgTerritory <- territoryPlot(vesalius, randomise = TRUE,cex =15 , cex.pt=3.5)



print(imgTerritory)



```  

# Territory extraction and markers

Finally, we can extract territories to run a more in depth analysis. This returns
a Seurat object with all selected territories. We can also find differential gene
expression between territories.


```{r extract, eval = TRUE, echo=TRUE}
# extract terriotry from seurat object
ter <- extractTerritories(vesalius,copy,seedID= 1)
ter

# get all markers related to all territories
all <- extractAllMarkers(vesalius,copy)
head(all)

# get markers associated with on or more territories
one <- extractMarkers(vesalius, copy, seed = c(1))
head(one)

# comparing territories
two <- extractMarkers(vesalius, copy, seed = c(1), query = c(2,3))
head(two)
```
