---
title: "Vesalius: Quick Start"
author: "Patrick C.N. Martin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vesalius: Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Vesalius: Quick Start

Do you like to live in the fast lane? Do you sigh every time you see a long
vignette? Do you wish there was a TLDR on everything?

I understand.

Welcome to the Vesalius quick start!

In this vignette, you will find a **very** short guide on how to work with Vesalius.
Vesalius provides an internal data set taken from real Spatial transcrtiptomic
data. 

We show *in context* Vesalius workflows in the articles section. 

For all intents and purposes, this is a dummy data set and should only be used
to get a feel for the Vesalius package. We strongly advise you to use real and
complete data sets to fully understand the Vesalius package and its benefits.


# Introduction
Vesalius is a tool to perform high-resolution in silico anatomization and
molecular characterization from Spatial omices data without requiring 
companion images. The nature of the vesalius algorithm makes it condusive to
being using on various spatial omics modalities without requiuring substantial 
changes. 

Vesalius achieves this by converting reducded dimensionality latent space into
gray scale images (one of each dimension). These images are processed using 
various image analysis techniques such as smoothing and segmentation.


# Loading & building a vesalius assay
## Loading data 

First, let's load the package and the data. Data originates from slide-seqV2
data (availble at [Single Cell Portal](https://singlecell.broadinstitute.org/single_cell/study/SCP948/robust-decomposition-of-cell-type-mixtures-in-spatial-transcriptomics#study-download)).

We took a small subset of the mouse hippocampus centered around the Dentate Gyrus. 

Here, we show an example using transcriptomic data but the same principles 
will apply for other modalities. For examples using other modalities, 
please refer to the other articles.

```{r loading, eval = TRUE, echo = TRUE}
suppressMessages(library(vesalius))
library(patchwork)
library(ggplot2)
data(vesalius, package = "vesalius")

```
### What do we have?
Once loaded we have two new variables: counts and cooridnates. 

The counts are the gene counts for all genes. As per convention, columns r
epresent the spatial indeces (barcodes, spots, beads,...) and rows are 
the genes that were captured. 

```{r str_counts, eval = TRUE, echo = TRUE}
str(counts)
```

The coordinates represent the x and y coordinates for each spatial index. 
```{r str_coordinates, eval = TRUE, echo = TRUE}
str(coordinates)
```

In other modalities, count matrices are still required for vesalius to run.

For example, Spatial-ATAC-seq or Spatial-cut&tag can be converted into count
matrices (more specifically tile matrices) using the [ArchR package](https://www.archrproject.com/)

## Building a Vesalius assay

We can build a vesalius assay by simply parsing both counts and 
cooridnates to the (vesalius assay contructor)[link to manual page].

```{r build_vesalius_arrary, eval = TRUE, echo = TRUE}
vesalius <- build_vesalius_assay(
  coordinates = coordinates, # spatial coordinates
  counts  = counts, # count matrix
  assay = "spatial_omics", # name you wish to give your assay
  adjust_coordinates = "origin", # origin snaps coordinates to origin (1,1)
  verbose = FALSE # Do you want progress messages?
)

```

### What do we have? 
A vesalius object is a container that requires at the very least spatial indices.
Here, we parsed both counts and spatial indices 

```{r show_ves, eval = TRUE, echo = TRUE}
vesalius
```

Count matrices can be added sperately using the `add_counts` function. 
For more information, please refere to the [Modifying vesalius objects]()

# Embedding latent space into gray scale images 
## Embed Latent Space

We can use this object top embed our latent space into grey scale images. 

```{r build_embeddings, eval = TRUE, echo = TRUE}
vesalius <- generate_embeddings(vesalius,
  dim_reduction = "PCA",
  normalisation = "log_norm",
  nfeatures = 100, # Setting number of features low for low run time
  verbose = FALSE)

```


### What do we have?
We see that we have added a set of embedding with one used one default embedding 
called the "active" embedding. See section below on *understanding active embeddings*.

```{r embedding_view, eval = TRUE, echo = FALSE}
vesalius
```

## Testing multiple embeddings 
We can test out multiple embeddings and every trial will be stored into 
the `vesalius_assay` object.
```{r adding_embeddings, eval = TRUE, echo = TRUE}
vesalius <- generate_embeddings(vesalius,
  dim_reduction = "UMAP",
  nfeatures = 100, # Setting number of features low for low run time
  verbose = FALSE)

vesalius <- generate_embeddings(vesalius,
  dim_reduction = "PCA",
  nfeatures = 200, # Setting number of features low for low run time
  verbose = FALSE)

vesalius <- generate_embeddings(vesalius,
  dim_reduction = "UMAP",
  nfeatures = 200, # Setting number of features low for low run time
  verbose = FALSE)
```
Every time you run a new dimensionality reduction approach it will be appended
to the previous ones. Vesalius does the same thing for different normalisation 
approaches. This lets you decided which combination you would want to use in 
later stages. 

Note that since you might be want to use the same embedding more than once, you will
end up with more than one embedding called "PCA" for example. In this case,
each embedding or normalisation will have a unique name. 

For every subsequent embedding called PCA, the trial name will be PCA.1, PCA.2 etc
The same applies to any embedding or normalisation method you select. 

### What do we have ?
```{r checkick_embed_names, eval = TRUE, echo = TRUE}
vesalius
```

You can also add your own embeddings and they will be appended into vesalius
in a similar manner. For more information, please refere to the [modifying vesalius objects]()
vignette. Alternively, you can also check the manual page for [add_embeddings]()


## Visualising Grey scale images

To visualise embeddings, we can simply plot the vesalius object and specify which 
dimension we wish to view. See (imagePlot)[link to page] for more information. 
Since we are leaving the embedding argument as the default "last" value, 
we can only look at the 3 UMAP dimensions. We can look at them simulaneously by
creating RGB images with one dimension per color channel. 
```{r grey_scale_view, eval = TRUE, echo = TRUE}
p1 <- image_plot(vesalius, dimensions = 1) + labs(title = "Grey UMAP dim 1")
p2 <- image_plot(vesalius, dimensions = seq(1, 3)) + labs(title = "RGB UMAP")
```
If we wanted to look at "PCA" instead, we can simply specify the embedding. 
Since we used 30 dimensions for PCA, we can use any dimensions to build RGB
images. 

```{r grey_scale_view_PCA, eval = TRUE, echo = TRUE}
p3 <- image_plot(vesalius, dimensions = 1, embedding = "PCA") +
  labs(title = "Grey PCA dim 1")
p4 <- image_plot(vesalius, dimensions = c(1, 4, 25), embedding = "PCA") +
  labs(title = "Grey PCA dim 1, 4, and 25")
```

```{r out_plot_grey, eval = TRUE, echo = TRUE, fig.width = 10, fig.height = 10}
(p1 + p2) / (p3 + p4)
```

# Understanding Active Embeddings

The active embedding slot in a `vesalius_assay` object contains the embedding values 
that will be used during image processing and image segmentation. 

In the examples above, we tested multiple embeddings.
```{r acive_embeds, eval = TRUE, echo = TRUE}
get_active_embedding_tag(vesalius)
```

We see that "UMAP.1" is used as the active embedding. Any time the embeddings are being 
used in subsequent functions they will be taken from here as this is the "last" embedding
used. 

In fact, in any subsequent function that contains the keyword "last" for one of its arguments,
vesalius will use the last target created. 
This can mean the last embedding used, the last count matrix 
or the last segmentation / territory isolation / territory morphology used.

If you are unstaified with the results, you can always call the embedding or trial you want 
to use explictely and the active slot will be replace with a fresh version of your embeddings
of choice. No need to re-run everything from scratch!

While this flexibility can be a boon, it can also be a curse! Anytime you explictely request 
an embedding or a trial it will use that fresh instance. This means that every time the active slot
will be updated and any processing you may have done to it will be overwritten. 




# Image processing and Image segmentation
## Image Processing

Once we have these images, we can apply image processing techniques to each gray scale image. 

```{r image_processing, eval = TRUE, echo = TRUE }
vesalius <- regularise_image(vesalius, verbose = FALSE)
vesalius <- equalize_image(vesalius, verbose = FALSE)
vesalius <- smooth_image(vesalius, iter = 10, sigma = 1, verbose = FALSE)

```

You can apply any image processing method you wish and in any order you desire. Please
note that the default number of dimensions for these function is defined by `seq(1,3)`
giving the first three dimensions. If working with PCA, please set this argument 
according to how many dimensions you wish to process. 

This also alows for the selction of abitrary of PCs. For example, based on the grey scale 
images, you might decided that this would be a better choice of PCs 

```{r selecting_PCS, eval = FALSE, echo = TRUE}
# Selecting a subset of PCs
dims <- c(1, 3, 4, 5, 7:11)

# running smoothing o
vesalius <- regularise_image(vesalius,
  dimensions = dims,
  embedding = "PCA",
  verbose = FALSE)
vesalius <- equalize_image(vesalius,
  dimensions = dims,
  verbose = FALSE)
vesalius <- smooth_image(vesalius,
  dimensions = dims,
  iter = 10,
  sigma = 1,
  verbose = FALSE)

```
In this example, we explicetely requested the *PCA* embedding. This means that we will take 
a fresh instance of the *PCA* embedding and use this as our active embedding. Once specified,
all subsequent functions will use this new active embedding unless specified otherwise.

However, note that dimensions are being parsed at every step. Only the images coressponding to 
these dimensions will be processed. Vesalius still retains the other ones just in case. 

## Image segmentation

Vesalius will attempts to segment the images into color segments. 
Vesalius use a kmeans clustering approach  to segment grey scale images. 
Note that the segmentation here is applied to the whole stack and not individually. 

The goal is top obtain colour segments that we can subdivived into territories. 
```{r segment_image, eval = TRUE, echo = TRUE}
vesalius <- segment_image(vesalius,
  method = "kmeans",
  col_resolution = 3,
  verbose = FALSE)
```

## What do we have?

We can see that the vesalius_assay object now contains some more information 
related to the segmentation.

```{r seg_check, eval = TRUE, echo = TRUE}
vesalius
```

We can also have a look at the results of the image segmentation.

```{r plot_segments_only, eval = TRUE, echo = TRUE, fig.width = 10, fig.height = 8}
p5 <- image_plot(vesalius) + labs(title = "Segments only")
print(p5)
```

Kmeans is the default setting for image segmentation but Vesalius also 
provides *louvain* and *leiden* based approaches. 

# Isolating territories
## Isolating territories from colour segments 

The final step is to isolate color segments into seperate territories. 
Similar color segments may be of the same colour but are seperated in 2D space. 
We want to be able to isolate each patch. For this, we can use the `isolate_territories`
function. 

```{r iso_territories, eval = TRUE, echo = TRUE}
vesalius <- isolate_territories(vesalius, capture_radius = 0.05)
```

### What do we have?
Now we can have a look at the isolated territories. We can keep track 
of how many instances of image segmentation and image manipulation we have gone through. 

```{r vesalius_with_ter, eval = TRUE, echo = TRUE}
vesalius
```



## Visualizing Territories 
We can plot our territories using vesalius plotting functions. Note that this is 
a `ggplot` object as such it is customisable using `ggplot2` functionalities. 
```{r plot_territories, eval = TRUE, echo = TRUE, fig.width = 10, fig.height = 8}
p6 <- territory_plot(vesalius, cex_pt = 3.5)
p6
```

# Comparing territories
Once we have our territories, we can compare the expression of genes between territories. 

```{r compare_territories, eval = TRUE, echo = TRUE}
vesalius <- identify_markers(vesalius, seed = 3, query = 5)
deg <- get_markers(vesalius)
```

And finally we can visualise the expression of our genes of interest. You can parse more than one gene.
Here we show the overall expression profile of **Malat1** but also its mean expression 
in each territory. 
```{r view_DEG, eval = TRUE, echo = TRUE, fig.width = 10,fig.height = 4}

p7 <- view_gene_expression(vesalius, genes = "Malat1")
p8 <- view_gene_expression(vesalius, genes = "Malat1", as_layer = TRUE)
p7 + p8
```

# See Also 

For more information about different aspects of Vesalius:

- [Modifying objects]()
- [Understanding vesalius_assay object]()

# Session Info
```{r, session, eval = TRUE, echo = TRUE}
sessionInfo()
```